---
title: "Visualization Project on Agricultural Crop Production in India"
author: 'Name: Meghna Mondal      Roll No. : MDS202123'
date: "29/11/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Introduction:

India’s agriculture is composed of many crops, with the foremost food staples being rice and wheat.
Indian farmers also grow pulses, potatoes, sugarcane, oilseeds, and such non-food items as cotton, tea, coffee, rubber, and jute.
Despite the overwhelming size of the agricultural sector, however, yields per hectare of crops in India are generally low compared to international standards.
So, here we wish to analyse more about the crop cultivation, cost invested in cultivation in different parts of our country.
Now we consider 2 datasets from the Kaggle dataset on Agricultural Crop Production in India (source:  https://www.kaggle.com/srinivas1/agricuture-crops-production-in-india/version/1?select=datafile+%282%29.csv ) namely datafile (1).csv and datafile (2).csv containing different variables described below.

```{r}

Variable_name = c("Crop", "State", "Cost.of.Cultivation....Hectare..A2.FL", "Cost.of.Cultivation....Hectare..C2", "Cost.of.Production....Quintal..C2", "Yield..Quintal..Hectare.")
Variable_type = c("Nominal", "Nominal", "Continuous", "Continuous", "Continuous", "Continuous")
Variable_content = c("Different types of crops", "States where the crop is cultivated", "Expected cost of cultivation of the crop per Hectare","Cost of cultivation of the crop per Hectare","Cost of Production per Quintal","Yield of crop in Quintal/Hectare")
Table_1 = data.frame( Variable_name, Variable_type, Variable_content)                         
knitr::kable(Table_1,"pipe",col.names = c("Variable Name","Variable Type","Variable Content"), align = c("l","c","c"))  



Variable_Name = c("Crop", "Production.2006.07", "Production.2007.08", "Production.2008.09", "Production.2009.10", "Production.2010.11", "Area.2006.07", "Area.2007.08", "Area.2008.09", "Area.2009.10", "Area.2010.11", "Yield.2006.07", "Yield.2007.08", "Yield.2008.09", "Yield.2009.10", "Yield.2010.11")
Variable_Type = c("Nominal", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous")  
Variable_Content = c("Different types of crops cultivated", "Production of a particular crop in Year 2006-07", "Production of a particular crop in Year 2007-08","Production of a particular crop in Year 2008-09","Production of a particular crop in Year 2009-10","Production of a particular crop in Year 2010-11"
                          , "Area of cultivation of particular crop in Year 2006-07", "Area of cultivation of particular crop in Year 2007-08", "Area of cultivation of particular crop in Year 2008-09", "Area of cultivation of particular crop in Year 2009-10", "Area of cultivation of particular crop in Year 2010-11"
                          , "Yield of a particular crop in Year 2006-07", "Yield of a particular crop in Year 2007-08","Yield of a particular crop in Year 2008-09","Yield of a particular crop in Year 2009-10","Yield of a particular crop in Year 2010-11")


Table_2 = data.frame( Variable_Name, Variable_Type, Variable_Content)                

knitr::kable(Table_2,"pipe",col.names = c("Variable Name","Variable Type","Variable Content"), align = c("l","c","c"))

```
In datafile(1) we have production cost, cultivation cost and yield of some specified crops in particular states. We have a series of data for a nominal attribute. So, we will use Bar Diagrams for analyzing those Cross-Sectional Data.

In datafile(2) we amount of production, area of cultivation and yield of some specified crops for five years. We have a series of data for a nominal attribute for varying time. So, we will use Bar Diagrams for analyzing those Panel Data.

```{r}

library(ggplot2)
setwd("C:/Users/MONDAL/Desktop/CMI/VISU")
crop_prod<-read.csv("datafile (2).csv")
crop_cost<-read.csv("datafile (1).csv")


ggplot(crop_cost,aes(y=Yield..Quintal..Hectare.,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in Quintal/Hectare")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Cropwise Yield")
ggplot(crop_cost,aes(y=Yield..Quintal..Hectare.,x=State))+geom_bar(stat="identity",aes(fill=State))+ylab("Yield in Quintal")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Statewise Yield")
ggplot(crop_cost,aes(y=Cost.of.Cultivation....Hectare..C2/100000,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Cost of Cultivation/Hectare in Lakh")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Cropwise Cost of cultivation per Hectare")
ggplot(crop_cost,aes(y=Cost.of.Cultivation....Hectare..C2/100000,x=State))+geom_bar(stat="identity",aes(fill=State))+ylab("Cost of Cultivation/Hectare in Lakh")+xlab("State")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Statewise Cost of cultivation per Hectare")
ggplot(crop_cost,aes(y=Cost.of.Cultivation....Hectare..A2.FL/100000,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Cost of Cultivation/Hectare in Lakh")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Cropwise Cost of cultivation per Hectare")
ggplot(crop_cost,aes(y=Cost.of.Cultivation....Hectare..A2.FL/100000,x=State))+geom_bar(stat="identity",aes(fill=State))+ylab("Cost of Cultivation/Hectare in Lakh")+xlab("State")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Statewise Cost of cultivation per Hectare")
ggplot(crop_cost,aes(y=Cost.of.Production....Quintal..C2/1000,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Cost of Production/Quintal in thousand")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Cost of production per Quintal")
ggplot(crop_cost,aes(y=Cost.of.Production....Quintal..C2/1000,x=Crop	,color=State))+geom_point(size=3)+ylab("Cost of Production/Quintal in Lakh")+xlab("Crop")+theme(axis.text.x=element_text(angle=90))+ggtitle("Statewise cost of production/Quintal")




ggplot(crop_cost,aes(y=Yield..Quintal..Hectare.,x=State,color=State))+geom_point(size=1)+ylab("Yied in Quintal")+xlab("Crop")+theme(axis.text.x=element_text(angle=90))+ggtitle("Statewise yield of crops/Hectare")+facet_wrap(~Crop	)
ggplot(crop_cost,aes(y=Cost.of.Cultivation....Hectare..C2,x=State,color=State))+geom_point(size=1)+ylab("Cost of cultivation")+xlab("Crop")+theme(axis.text.x=element_text(angle=90))+ggtitle("Statewise Cost of Cultivation/Hectare")+facet_wrap(~Crop	)
ggplot(crop_cost,aes(y=Cost.of.Production....Quintal..C2,x=State,color=State))+geom_point(size=1)+ylab("Cost of production")+xlab("Crop")+theme(axis.text.x=element_text(angle=90))+ggtitle("Statewise Cost of Production/Quintal")+facet_wrap(~Crop	)




options(repr.plot.width=15,repr.plot.height=10)
pairs(crop_cost[,-c(1:2)])
cor(crop_cost[,-c(1:2)])




ggplot(crop_prod,aes(y=Production.2006.07,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Production in 2006-07")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Production in 2006-07")
ggplot(crop_prod,aes(y=Production.2007.08,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Production in 2007-08")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Production in 2007-08")
ggplot(crop_prod,aes(y=Production.2008.09,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Production in 2008-09")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Production in 2008-09")
ggplot(crop_prod,aes(y=Production.2009.10,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Production in 2009-10")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Production in 2009-10")
ggplot(crop_prod,aes(y=Production.2010.11,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Production in 2010-11")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Production in 2010-11")


ggplot(crop_prod,aes(y=Area.2006.07,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Area in 2006-07")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Area in 2006-07")
ggplot(crop_prod,aes(y=Area.2007.08,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Area in 2007-08")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Area in 2007-08")
ggplot(crop_prod,aes(y=Area.2008.09,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Area in 2008-09")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Area in 2008-09")
ggplot(crop_prod,aes(y=Area.2009.10,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Area in 2009-10")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Area in 2009-10")
ggplot(crop_prod,aes(y=Area.2010.11,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Area in 2010-11")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Area in 2010-11")


ggplot(crop_prod,aes(y=Yield.2006.07,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in 2006-07")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Yield in 2006-07")
ggplot(crop_prod,aes(y=Yield.2007.08,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in 2007-08")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Yield in 2007-08")
ggplot(crop_prod,aes(y=Yield.2008.09,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in 2008-09")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Yield in 2008-09")
ggplot(crop_prod,aes(y=Yield.2009.10,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in 2009-10")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Yield in 2009-10")
ggplot(crop_prod,aes(y=Yield.2010.11,x=Crop	,fill=Crop	))+geom_bar(stat="identity")+ylab("Yield in 2010-11")+xlab("Crop")+theme(axis.text.x=element_text(angle=90),legend.position="none")+ggtitle("Yield in 2010-11")





Production = data.frame(crop_prod$Production.2006.07,crop_prod$Production.2007.08,crop_prod$Production.2008.09,crop_prod$Production.2009.10,crop_prod$Production.2010.11)
P = as.matrix(Production)
par(mfrow=c(1,1))
barplot(t(P), main="Multiple Bar Diagram of Production in five year for different crop", ylab="Total", beside=TRUE, 
        col=terrain.colors(55))
options(repr.plot.width=15,repr.plot.height=10)
pairs(Production)
cor(Production)

Area = data.frame(crop_prod$Area.2006.07,crop_prod$Area.2007.08,crop_prod$Area.2008.09,crop_prod$Area.2009.10,crop_prod$Area.2010.11)
A = as.matrix(Area)
par(mfrow=c(1,1))
barplot(t(A), main="Multiple Bar Diagram of Production in five year for different crop", ylab="Total", beside=TRUE, 
        col=terrain.colors(55))
options(repr.plot.width=15,repr.plot.height=10)
pairs(Area)
cor(Area)


Yield = data.frame(crop_prod$Yield.2006.07,crop_prod$Yield.2007.08,crop_prod$Yield.2008.09,crop_prod$Yield.2009.10,crop_prod$Yield.2010.11)
Y = as.matrix(Yield)
par(mfrow=c(1,1))
barplot(t(Y), main="Multiple Bar Diagram of Production in five year for different crop", ylab="Total", beside=TRUE, 
        col=terrain.colors(55))
options(repr.plot.width=15,repr.plot.height=10)
pairs(Yield)
cor(Yield)

Year_2006_07 = data.frame(crop_prod$Production.2006.07,crop_prod$Area.2006.07,crop_prod$Yield.2006.07)

pairs(Year_2006_07)
cor(Year_2006_07)

Year_2007_08 = data.frame(crop_prod$Production.2007.08,crop_prod$Area.2007.08,crop_prod$Yield.2007.08)
pairs(Year_2007_08)
cor(Year_2007_08)

Year_2008_09 = data.frame(crop_prod$Production.2008.09,crop_prod$Area.2008.09,crop_prod$Yield.2008.09)
pairs(Year_2008_09)
cor(Year_2008_09)

Year_2009_10 = data.frame(crop_prod$Production.2009.10,crop_prod$Area.2009.10,crop_prod$Yield.2009.10)
pairs(Year_2009_10)
cor(Year_2009_10)

Year_2010_11 = data.frame(crop_prod$Production.2010.11,crop_prod$Area.2010.11,crop_prod$Yield.2010.11)
pairs(Year_2010_11)
cor(Year_2010_11)




```


Conclusion:

              From the univariate and bivariate analysis of the above dataset we get an idea about in which state, how much amount of crop are cultivated in how much area.  
For e.g., we get to know the yield and cost of cultivation per Hectare are comparatively higher for sugarcane i.e., it is being produced in a large amount. 
 Again, in Andhra Pradesh, Maharashtra, Karnataka, Tamil Nadu, Uttar Pradesh yield produced are much higher than the other states.
In the context of cost of cultivation Andhra Pradesh shows much higher values than the other states. 
Eventually we can infer this type of information from each graph. 
We also have computed the correlation matrix for the variables considered. If the correlation coefficient comes out to be near 1 two variables said to be positively correlated, or if the correlation coefficient comes out to be near -1 two variables said to be negatively correlated and if it comes out to be near 0, we can expect the two variables to be independent on each other.
For e.g., we observe for all the five years the crop production and yield are highly positively correlated but the area and yield are area moderately related.
For easy visual interpretation we also plotted the scatter diagram of these variables.
